Class-based network segregation, Economic Inequality and Redistributive Preferences across societies

### 2025-07-18 paper1_jiturra.Rmd
RR Patrick

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  prefineq,redist,
  egal = egal2,
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  # religion,
  # union,
  workst, 
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN")

@@

digclas1_VW_7 <- update(digclas1_VW_3, . ~ . +dclass3res_V*homclass3_V_res)
digclas1_VW <- list(digclas1_VW_0,digclas1_VW_1,
                    # digclas1_VW_3,
                    # digclas1_VW_4,
                    #digclas1_VW_5,
                    # digclas1_VW_6,
                    digclas1_VW_7)

@@

texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),single.row = T,symbol = ".",
       caption = paste("(\\#tab:table1)","Multilevel models for network homogeneity and redistributive preferences"),
       custom.note = "\\item Note: Models include sampling weights. \\item Standard errors in parentheses. %stars",threeparttable = T,
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,scalebox = 0.7,
      groups = list("Social Class (Ref.= Service Class)" = 3:4, 
      "Homogeneity x Social Class"=5:6),
      custom.gof.rows = list("Controls"=c("No",rep("Yes",2))),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)

@@

```{r fig4, fig.dim=c(10,5),fig.cap='Interaction of network homogeneity and social class on redistributive preferences',warning=FALSE}
int_homo <- digclas1_VW_7
df_pred1 <- 
  predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"),
              newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res), max(dfreg$homclass3_V_res), by= 0.1),
                                 dclass3res_V = levels(dfreg$dclass3res_V)))
# names(df_pred1) 
df_pred1 %>% 
ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=2,color="black") +
  facet_grid(~factor(dclass3res_V, levels=
                       c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"),
                     c("Working Class","Intermediate Class","Service Class"))) +
  scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_y_continuous(limits = c(55,80))+
  labs(caption = paste0("Note: N (individuals) = ",nobs(int_homo),"; N (countries) = ",lme4::ngrps(int_homo),". Predictive estimates with 95% confidence intervals"),
       x="Network homogeneity",y="Redistributive preferences (0 - 100)") +
  theme(legend.position = "none", legend.title = element_blank())

@@

dfreg <- df1 %>% dplyr::select(
  redist,prefineq,
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  # dclass3spo_V, 
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  # religion,
  partner,
  # union,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  # "gini_disp"=difclass,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

dfreg$wstate2 <- scale(scale(dfreg$abs_red) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))

@@

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2,homo_V_gini_full3)
screenreg(digclas1_macro)

@@

ggsave(plot = arrangeGrob(fig_homodiff_gini, nrow = 1),filename = "slides-bivar-homodiff.png",device = "png",
       path = here::here("output/images"),width = 1,height = 1,units = "cm",scale = 17)

>>>>

load(here::here("input/data/proc/study1_country.RData"));df1 <- df2
dfreg <- df1 %>% dplyr::select(
  prefineq,redist,
  egal = egal2,
  dclass3res_V,
  # dclass3spo_V,
  homclass3_V_res,
  know_total,
  # socialtrust,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  # religion,
  # union,
  workst, 
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  mutate(logrgdpna = log(rgdpna),
         loggdppercapita=log(gdppercapita),
         edyears2 = edyears ^ 2,
         age2 = agenum ^ 2) %>%
  # filter(oecd == "OECD") %>%
  filter(country2 != "SVN") 
dfreg$egal <- as.numeric(scale(dfreg$egal))

@@

digclas1_VW_7 <- update(digclas1_VW_3, . ~ . +dclass3res_V*homclass3_V_res)
digclas1_VW <- list(digclas1_VW_0,digclas1_VW_1,
                    digclas1_VW_3,
                    # digclas1_VW_4,
                    #digclas1_VW_5,
                    # digclas1_VW_6,
                    digclas1_VW_7)

@@

screenreg(digclas1_VW)
texreg(digclas1_VW,stars = c(0.001, 0.01, 0.05, 0.1),single.row = T,symbol = ".",
       caption = paste("(\\#tab:table1)","Multilevel models for network homogeneity and redistributive preferences"),
       custom.note = "\\item Note: Models include sampling weights. \\item Standard errors in parentheses. %stars",threeparttable = T,
       caption.above = T,leading.zero = T,
       custom.coef.map = ccoef,scalebox = 0.7,
      groups = list("Social Class (Ref.= Service Class)" = 3:4, 
      "Homogeneity x Social Class"=5:6),
      custom.gof.rows = list("Controls"=c("No",rep("Yes",2))),include.loglik = FALSE,include.aic = FALSE,
      custom.gof.names = c(NA,NA,"Num. groups","Var: Country (Intercept)","Var: Residual"),
       float.pos = "h!",
       use.packages = F)

@@

```{r fig4, fig.dim=c(10,5),fig.cap='Interaction of network homogeneity and social class on redistributive preferences',warning=FALSE}
int_homo <- digclas1_VW_7
df_pred1 <- 
  predictions(int_homo, condition = c("homclass3_V_res","dclass3res_V"),
              newdata = datagrid(homclass3_V_res = seq(min(dfreg$homclass3_V_res),
                                                       max(dfreg$homclass3_V_res), by= 0.2),
                                 dclass3res_V = levels(dfreg$dclass3res_V),
                                 know_total = mean(dfreg$know_total)))
# names(df_pred1) 
df_pred1 %>% 
ggplot(aes(y =estimate,x=homclass3_V_res, fill=dclass3res_V, color=dclass3res_V,ymin=conf.low, ymax=conf.high)) +
  geom_ribbon(alpha=0.8,size=1,linetype="solid",color="black") +
  geom_line(size=1,color="black") +
  geom_point(size=2,color="black") +
  facet_grid(~factor(dclass3res_V, levels=
                       c("Working Class (V+VI+VII)","Intermediate class (III+IV)","Service Class (I+II)"),
                     c("Working Class","Intermediate Class","Service Class"))) +
  scale_fill_manual(values=c("#323232","#989898","#CCCCCC"))+
  scale_color_manual(values=c("#323232","#989898","#CCCCCC"))+
  # scale_y_continuous(limits = c(55,80))+
  labs(caption = paste0("Note: N (individuals) = ",nobs(int_homo),"; N (countries) = ",lme4::ngrps(int_homo),". Predictive estimates with 95% confidence intervals"),
       x="Network homogeneity (z-score)",y="Redistributive preferences (z-score)") +
  theme(legend.position = "none", legend.title = element_blank())

@@

dfreg <- df1 %>% dplyr::select(
  redist,prefineq,
  egal = egal2,
  # starts_with("class3"),
  # starts_with("digclass3"),
  # starts_with("dclass3"),
  # starts_with("homclass"),
  dclass3res_V,
  homclass3_V_res,
  know_total,
  Q03pcm,
  edyears,
  female,
  agenum,
  partner,
  workst,
  WEIGHT,
  region,
  "gini_disp"=wid_gini_disp,
  "gini_mkt",
  gv_spen,
  rel_red,
  abs_red,
  ilo_taxrev,
  ilo_govspe,
  d10d1=wid_p90p10,
  wid_p90p50,
  top10=wid_sharetop10,
  rgdpna,
  gdppercapita,
  oecd,
  country2, country, country3,
) %>%
  filter(country2 != "SVN") %>% 
  na.omit() %>% 
    mutate(loggdppercapita=scale(gdppercapita/1000),
         age2 = agenum ^ 2, 
         gini_disp=scale(gini_disp))

dfreg$wstate2 <- scale(scale(dfreg$abs_red) + scale(dfreg$ilo_taxrev) + scale(dfreg$ilo_govspe))
dfreg$egal <- as.numeric(scale(dfreg$egal))
# dfreg$homclass3_V_res <- as.numeric(scale(dfreg$homclass3_V_res))
# dfreg$know_total <- as.numeric(scale(dfreg$know_total))
# dfreg$agenum <- as.numeric(scale(dfreg$agenum))
# dfreg$edyears <- as.numeric(scale(dfreg$edyears))

@@

digclas1_macro <- list(homo_V_gini,
                       homo_V_gini_gdp,
                       homo_V_gini_gdp_wstate2,
                       homo_V_gini_full2,homo_V_gini_full3)

screenreg(digclas1_macro)

@@

ggsave(plot = arrangeGrob(fig_homodiff_gini, nrow = 1),filename = "slides-bivar-homodiff.png",device = "png",
       path = here::here("output/images"),width = 1,height = 1,units = "cm",scale = 17)
```
